Appendix to 'Detecting social information in a dense database of infants’ natural

### 2021-06-09 paper/preprint/appendix.Rmd

# Face/hand detections relative to human annotations
```{r goldSetSanity, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces and hands seen as a function of age for each child in the dataset. Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range. Dashed lines show estimated trend lines from proportion of faces/hands in view when analyzing the gold set of frames made by human annotators. Dotted lines show trend lines from the goldset when frames when children's own hand were detected." }

ggplot(all_detections %>% filter(child_id %in% c('S','A')), 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = FALSE) + 
  geom_smooth(data = goldset_adult_hands, span=10, aes(weight = num_frames_total), show.legend = FALSE,
              lty = 2, span=10, se = FALSE) +
  geom_smooth(data = goldset, span=10, aes(weight = num_frames_total), show.legend = FALSE,
              lty = 3, span=10, se = FALSE) +
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.6) +
  facet_grid(.~child_id) + 
  theme_few(base_size=9) +
  ggthemes::scale_color_solarized(name = "") + 
  scale_size_continuous(name = "Detections (Log 10)") +
  theme(legend.text=element_text(size=8)) +
  theme(legend.position="bottom")

@@

```{r}
### plot centers of the bounding boxes made for child and adult hands
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ggtitle('A. Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=6) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=6) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ggtitle('B. Adult hand density')
```

```{r handLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=5, fig.height=3, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant. Brighter values indicate more detections."}
cowplot::plot_grid(child_hand_plot, adult_hand_plot, nrow=1)
```

# Distribution of faces and hands in the visual field
We explored where in the visual field children tended to see faces and hands, suspecting that these distributions might become wider as children grow older and learn to locomote on their own, following preliminary analyses from Frank (2012). As expected, faces tended to appear in the upper visual field in contrast to hands, which tended to be more centrally located.However, we found little evidence for any changes in the positions of faces and hands across age, suggesting that this is a relatively stable property of infants’ visual environment from 6 months of age.

>>>>

# Face/hand detections relative to human annotations
```{r goldSetSanity, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion of faces and hands seen as a function of age for each child in the dataset. Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range. Dashed lines show estimated trend lines from proportion of faces/hands in view when analyzing the gold set of frames made by human annotators. Dotted lines show trend lines from the goldset when frames when children's own hand were detected." }

ggplot(all_detections %>% filter(child_id %in% c('S','A')), 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = FALSE) + 
  geom_smooth(data = goldset_adult_hands, span=10, aes(weight = num_frames_total), show.legend = FALSE,
              lty = 2, span=10, se = FALSE) +
  geom_smooth(data = goldset, span=10, aes(weight = num_frames_total), show.legend = FALSE,
              lty = 3, span=10, se = FALSE) +
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.6) +
  facet_grid(.~child_id) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  scale_size_continuous(name = "Detections (Log 10)") +
  theme(legend.text=element_text(size=10)) +
  theme(legend.position="bottom")

@@

```{r}
### plot centers of the bounding boxes made for child and adult hands
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ggtitle('A. Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=10) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=10) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ggtitle('B. Adult hand density')
```

```{r handLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=6.5, fig.height=3.5, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant. Brighter values indicate more detections."}
cowplot::plot_grid(child_hand_plot, adult_hand_plot, nrow=1)
```

# Distribution of faces and hands in the visual field
We explored where in the visual field children tended to see faces and hands, suspecting that these distributions might become wider as children grow older and learn to locomote on their own, following preliminary analyses from Frank (2012). 
As expected, faces tended to appear in the upper visual field in contrast to hands, which tended to be more centrally located (see Figure C1). 
However, we found little evidence for any changes in the positions of faces and hands across age, suggesting that this is a relatively stable property of infants’ visual environment from 6 months of age.

### 2021-06-09 paper/preprint/paper_revised_submission_april2021_apa.Rmd

keywords          : "social cognition; face perception; infancy; head cameras; deep learning"
wordcount         : "4671"

@@

With hundreds of hours of footage (>40M frames), however, this large dataset necessitates a shift to an automated annotation strategy.
Indeed, annotation of the frames extracted from egocentric videos has been prohibitively time-consuming, meaning that most frames are typically not inspected, even in the most comprehensive studies. 
For example, @Fausey2016 collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand-annotated (by four coders), totalling 103,383 frames (per coder)—an impressive number of annotations but nonetheless only 0.67% of the collected footage.
To address this challenge, we use a modern computer vision model of pose detection to automatically detect the presence of hands and faces from the infant egocentric viewpoint. 
Specifically, we use OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints that operates well on scenes including multiple people, even if they are partially-occluded (see Figure 1). 
In prior work examining egocentric videos, OpenPose performed comparably to other modern face detection models [@sanchez2018postural].

@@

## Dataset
The dataset is described in detail in @SAYcam; we summarize these details here. Children wore Veho Muvi miniature cameras mounted on a custom camping headlamp harness ("headcams") at least twice weekly, for approximately one hour per recording session. One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family’s discretion. At the time of the recording, all three children were in single-child households.  Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase the field of view to approximately 109 degrees horizontal x 70 degrees vertical. Videos ^[All videos are available at https://nyu.databrary.org/volume/564 ] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset.  We analyze `r num_videos` videos, with a total duration of `r round(num_hours,2)` hours (>40 million frames).

@@

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=6, fig.cap = "Proportion of frames with (A).All face and hand detections,  B.Face/hand detections that fell within the center field-of-view (reducing the contribution of children's own hands) and (C) Face detections that were full faces (e.g., eyes, nose, and mouth all visible) and that co-occurred with hands, plotted as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range."}

@@

We also found an unexpectedly high proportion of hands in the view of infants, even when restricting the field-of-view to the center field-of-view to make the viewpoints comparable to those of headcams used in prior work. Why might this be the case? One idea is that these videos contain the viewpoints of children not only during structured interactions (e.g., play sessions at home or in the lab) but during everyday activities when children may be playing by themselves or simply observing the actions of caregivers and other people in their environment. During these less structured times, caregivers may move about in the vicinity of the child but not interact with them as directly -- leading to views where a person and their hands are visible from a distance, but this person's face may be turned away from the infant or occluded (see examples in Figure 1). Indeed, using the same pose detector on videos from in-lab play sessions, @sanchez2018postural found the opposite trend: slightly fewer hand detections than face detections from 8-16 months of age. Work that directly examines the variability in the social information in view across more vs. less structured activity contexts could further test this idea.

@@

We also used these detailed pose annotations to explore finer-grained changes in how children experience the faces and hands of their caregivers over development. We found that the faces, hands, and people in the infant view tended to become smaller and that faces tended to be farther away and in more variable horizontal positions, in keeping with prior work examining the sizes of faces in the infant view during the first year of life [@Jayaraman2015]. Overall, these data support the idea that the social information in view changes across development as infants become increasingly mobile and independent [@franchak2018; @fausey2016]. As children explore the world on their own [@xu2019towards], they may experience fewer close-up interactions with the caregivers and more bouts of play where they are exploring the objects and things in the environment around them.

More broadly, however, these analyses underscore the importance of how, when, from whom, and what data we sample; these choices become central when we attempt to draw conclusions about the regularities of experience. Indeed, while unprecedented in size, this dataset still has many limitations. These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative in many ways of the global population [@karasik2018ties; @henrich2010most]. Any idiosyncrasies in how and when these particular families chose to film these videos also undoubtedly influences the variability seen here, and may contribute to the individual differences between the three children in this dataset. And without eye-tracking data, we do not know if children are attending to the social information in their visual field.

Nonetheless, we believe that these advances in datasets and methodologies represent a step in the right direction. The present paper demonstrates the feasibility of using a modern computer vision model to annotate the entirety of a very large dataset (here, >40M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to human annotations in prior work. While OpenPose did not provide annotations that were as accurate as those provided by human annotators, we found relatively consistent results with prior literature, suggesting that the sheer scale and density of the annotations provided by this method may overcome some of it's limitations.

>>>>

keywords          : "social cognition; face perception; infancy; head cameras; deep learning"
wordcount         : "4698"

@@

With hundreds of hours of footage (>42M frames), however, this large dataset necessitates a shift to an automated annotation strategy.
Indeed, annotation of the frames extracted from egocentric videos has been prohibitively time-consuming, meaning that most frames are typically not inspected, even in the most comprehensive studies. 
For example, @Fausey2016 collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand-annotated (by four coders), totalling 103,383 frames (per coder)—an impressive number of annotations but nonetheless only 0.67% of the collected footage.
To address this challenge, we use a modern computer vision model of pose detection to automatically detect the presence of hands and faces from the infant egocentric viewpoint. 
Specifically, we use OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints that operates well on scenes including multiple people, even if they are partially-occluded (see Figure 1). 
In prior work examining egocentric videos, OpenPose performed comparably to other modern face detection models [@sanchez2018postural].

@@

## Dataset
The dataset is described in detail in @SAYcam; we summarize these details here. Children wore Veho Muvi miniature cameras mounted on a custom camping headlamp harness ("headcams") at least twice weekly, for approximately one hour per recording session. One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family’s discretion. At the time of the recording, all three children were in single-child households.  Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase the field of view to approximately 109 degrees horizontal x 70 degrees vertical. Videos ^[All videos are available at https://nyu.databrary.org/volume/564 ] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset.  We analyze `r num_videos` videos, with a total duration of `r round(num_hours,2)` hours (>42 million frames).

@@

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=6, fig.cap = "Proportion of frames with (A) All face and hand detections,  (B) Face/hand detections that fell within the center field-of-view (reducing the contribution of children's own hands) and (C) Face detections that were full faces (e.g., eyes, nose, and mouth all visible) and that co-occurred with hands, plotted as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range."}

@@

We also found an unexpectedly high proportion of hands in infants' view, even when restricting the field-of-view to the center field-of-view to make the viewpoints comparable to those of headcams used in prior work. 
Why might this be the case? One idea is that these videos contain the viewpoints of children not only during structured interactions (e.g., play sessions at home or in the lab) but during everyday activities when children may be playing by themselves or simply observing the actions of caregivers and other people in their environment. 
During these less structured times, caregivers may move about in the vicinity of the child but not interact with them as directly -- leading to views where a person and their hands are visible from a distance, but this person's face may be turned away from the infant or occluded (see examples in Figure 1). Indeed, using the same pose detector on videos from in-lab play sessions, @sanchez2018postural found the opposite trend: slightly fewer hand detections than face detections from 8-16 months of age. Work that directly examines the variability in the social information in view across more vs. less structured activity contexts could further test this idea.

@@

We also used these detailed pose annotations to explore finer-grained changes in how children experience the faces and hands of their caregivers over development. We found that the faces, hands, and people in the infant view tended to become smaller and that faces tended to be farther away and in more variable horizontal positions, in keeping with prior work examining the sizes of faces in the infant view during the first year of life [@Jayaraman2015]. 
Overall, these data support the idea that the social information in view changes across development as infants become increasingly mobile and independent [@franchak2017see; @fausey2016]. 
As children explore the world on their own [@xu2019towards], they may experience fewer close-up interactions with their caregivers and more bouts of play where they are exploring the objects in their environment.

More broadly, however, these analyses underscore the importance of how, when, from whom, and what data we sample; these choices become central when we attempt to draw conclusions about the regularities of experience. 
Indeed, while unprecedented in size, this dataset still has many limitations. 
These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative in many ways of the global population [@karasik2018ties; @henrich2010most]. 
Any idiosyncrasies in how and when these particular families chose to film these videos also undoubtedly influenced the variability seen here, and may contribute to the individual differences between the three children in this dataset. 
And without eye-tracking data, we do not know the extent to which children are attending to the social information in view.

Nonetheless, we believe that these advances in datasets and methodologies represent a step in the right direction. 
The present paper demonstrates the feasibility of using a modern computer vision model to annotate the entirety of a very large dataset (here, >42M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to human annotations in prior work. 
While OpenPose did not provide annotations that were as accurate as those provided by human annotators, we found relatively consistent results with prior literature, suggesting that the sheer scale and density of the annotations provided by this method may overcome some of its limitations.
